
\textbf{I have used Generative AI in this work} \quad \fbox{YES}

\section*{Part 1: Reflection on ethical and responsible behavior}

\textbf{1. In my interaction with Generative AI tools, I have submitted sensitive data with the proper authorization of the interested parties.}
\begin{itemize}
  \item[] \fbox{NO, I have not used sensitive data}
\end{itemize}

\textbf{2. I have submitted copyrighted materials with the proper authorization.}
\begin{itemize}
  \item[] \fbox{NO, I have not used protected materials}
\end{itemize}

\textbf{3. I have submitted personal data with the proper authorization of the interested parties.}
\begin{itemize}
  \item[] \fbox{NO, I have not used personal data}
\end{itemize}

\textbf{4. My use of the tool has respected the terms of use and ethical principles.}
\begin{itemize}
  \item[] \fbox{YES}
\end{itemize}


\section*{Part 2: Technical Use Declaration}


\subsection*{Documentation and writing}
I have used Generative AI for:
\begin{itemize}
  \item Search of bibliography to \href{https://www.google.com}{Google Scholar}, \href{https://openai.com}{GPT o3, o4}.
  \item Generation of tables based on provided data.
\end{itemize}

\subsection*{Development of specific content}
\begin{itemize}
  \item Programming assistance. \\
  I have asked for explanations of Python, Go and C++ code to \href{https://openai.com}{GPT 4o Copilot}, \href{https://www.anthropic.com/api}{Claude (3.5, 3.7, 4)}, \href{https://deepmind.google/models/gemini/pro/}{Gemini 2.5 Pro} and \href{https://qwenlm.github.io/}{Qwen 2.5}.

  \item Generation of elements. \\
  I have asked for generation of diagrams and tables based on benchmark data to \href{https://openai.com}{OpenAI's GPT 4.1, 4o, o4}, \href{https://deepmind.google/models/gemini/pro/}{Gemini 2.5 Pro}, \href{https://qwenlm.github.io/}{Qwen3} and \href{https://www.anthropic.com/api}{Claude (3.5, 4)}.

  \item Read proofing and correction of texts. \\
  I have asked for corrections of the text in the appendix and the main body of the work to \href{https://openai.com}{Gemini 2.5 Pro}.
\end{itemize}

\section*{Part 3: Reflection on Utility}

The use of Generative AI tools has been beneficial in several ways. It has allowed me to quickly generate tables and diagrams based on benchmark data, which would have taken significantly more time to create manually and style consistently. Additionally, the ability to ask for explanations of code and receive immediate feedback has been invaluable in using new programming techniques.

I have not used AI to debug code as I have found out that by using AI to debug errors, it can lead to a lack of understanding of the underlying issues, and it can lead to an erroneous fix. 

In the process, I have learned also how LLMs work and what is the best way to use them, never accepting changes blindly, always checking the output.